---
title: "Lecture 7 Graphical Data Analysis"
author: "The Great Courses"
output: 
  beamer_presentation: 
    keep_tex: yes
    colortheme: "albatross"
    pandoc_args: [ 
      "--template", "./beamer_template.tex"
    ]
---
```{r setup, echo=FALSE, include=FALSE}
knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before) 
    return(options$size)
})
knitr::opts_chunk$set(mysize=TRUE)
knitr::opts_chunk$set(size='\\small')
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(fig.height=3)
knitr::opts_chunk$set(fig.width=4)
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(fig.align='center')
knitr::opts_chunk$set(comment="")
knitr::opts_chunk$set(collapse = TRUE)

# use of knit_hooks$set for par option to be applied to all the chunks
knitr::knit_hooks$set(mylightplot = function(before, options, envir) {
    if (before) 
      par(col="gray",      # color of lines
          col.axis="gray", # color of axis annotation
          col.lab="gray",  # color of x and y labels
          col.main="gray", # color of the main title
          col.sub="gray",  # color of the subtitles
          col.ticks="gray",# color of tick marks
          fg="lightgray")  # color of axes, boxes
})
knitr::opts_chunk$set(mylightplot=TRUE)

```

## Today's topic: Graphical Data analysis

Statistical graphs are useful in helping us visualize data.  Through graphs, we:

- Understand data properties
- Find patterns in data
- Suggest modeling strategies
- "Debug" our analyses
- Communicate results

## Learning Objectives for today

- Define and identify basic numerical and graphical summeries of data

- Use R for calculating descriptive statistics, making graphs, and writing functions

## Iris Data

The Iris dataset is widely used throughout statistical science for illustrating various problems in statistical graphics, multivariate statistics and machine learning.

- It's a small, but non-trivial dataset.

- The data values are real (as opposed to simulated) and are of high quality (collected with minimal error). 

- The data were used by the celebrated British statistician Ronald Fisher in 1936. (Later he was knighted and became Sir Ronald.) 

- Using a few famous datasets is one of the traditions we hand down in statistics!  (Also, when comparing old and new methods, or in evaluating any method, it's helpful to try them out on known datasets, thus maintaining continuity in how we assess methods.) 

## Iris Data
The Iris dataset is most commonly used for on pattern recognition in statistics.  The dataset contains 3 classes of 50 instances each, where each class refers to a type of iris plant, with the following attributes: 

- Sepal Length 
- Sepal Width 
- Petal Length  
- Petal Width 
- class: Iris setosa, Iris versicolor, Iris virginica

<!-- ## Iris Data features -->

<!-- ![Iris](iris_petal_sepal.bmp) -->

<!-- ## Iris Data features -->

<!-- The Iris Species are difficult to visually discern due to their similarity.  But are there statistical differences that might aid our discernment? -->

<!-- ![Iris Types](iris_types.bmp) -->

## Load Data

The Iris data is in the `datasets` library in R.  Type the following commands:

```{r}
library(datasets)
library(RColorBrewer)
attach(iris)
head(iris)
```

## Bar Plots
Let's begin our analysis.

- **Bar Plots** are useful for showing comparisons across several groups.  Although it looks like a histogram, a bar plot is plotted over a label that represents a category (e.g., Iris type).

- One indication of the difference between a bar plot and histogram: It's always appropriate to talk about the skewness of a histogram; that is, the tendency of the observations to fall more on the low end or the high end of the X axis.

- However, on bar plots, the X axis can sometimes be categorical - (i.e. not quantitative.) 

## Bar Plots
```{r barplot1}
barplot(iris$Petal.Length, main = "Petal Length") 
```

## Bar Plot
```{r barplot2}
barplot(iris$Sepal.Length, 
        col= brewer.pal(3,"Set1"), main = "Sepal Length")
```

## Bar Plot
```{r barplot3}
barplot(table(iris$Species,iris$Sepal.Length), 
        col  = brewer.pal(3,"Set1"), 
        main = "Stacked Plot of Sepal Length by Species")
```

## Summary Statistics
The `summary` function is a quick and easy way to assess the statistical properties of each attribute.  These values are displayed graphically in a box plot.

```{r}
summary(iris)
```

## Box Plots
- Box plots are used to compactly show many pieces of information about a variables distribution and is useful for visualizing the spread of the data.

- Box plots show **five statistically important numbers** - the minimum, the 25th percentile, the median, the 75th percentile and the maximum. 

```{r}
boxplot(iris$Sepal.Length, main = "Sepal Length")
```

\vfill

## Box Plots
```{r boxplot1}
boxplot(iris[,1:4], 
        names=c("Sep L", "Sep W", "Pet L", "Pet W"))
```

## Box Plots
A box plot can also be used to show how one attribute `petal.length` varies with another attribute `iris.type`.

```{r}
boxplot(iris$Petal.Length~iris$Species, 
        main = "Petal Length vs. Species") 
```

## Box Plot
We can visualize how the spread of `Sepal Length` changes across various categories of `Species`.  A color palette is a group of colors that is used to make the graph more appealing and help create visual distinctions in the data.

## Box Plot
```{r}
boxplot(iris$Sepal.Length~iris$Species,
        col=heat.colors(3), 
        main = "Sepal Length vs. Species")
```

## Scatter Plot

Scatter plots help in visualizing data easily and for simple data inspection.  Try the following code.

```{r}
plot(iris$Petal.Length, main="Petal Length", 
     ylab = "Petal Length", xlab = "Species") 
```

## Scatter Plot
Let's generate corresponding scatterplots for Petal.Width, Sepal.Length and Sepal.Width.  

```{r}
par(mfrow=c(1,2))
plot(iris$Petal.Length, main="Petal Length") 
plot(iris$Sepal.Length, main="Sepal Length")
```
## Scatter Plot
```{r}
par(mfrow=c(1,2))
plot(iris$Petal.Width, main="Petal Width") 
plot(iris$Sepal.Width, main="Sepal Width") 
```

What are our observations?  Which plots help us distinguish between groups?

## Scatter Plots

- Scatter plots are used to plot two variables against each other.  We can add a third dimension by coloring the data values according to their Species. 

- For datasets with only a few attributes, we can construct and view all the pairwise scatter plots.

## Pairwise Plots
```{r, eval=FALSE, echo=TRUE,results=FALSE}
pairs(as.matrix(iris[,-5]), pch=21, bg=c("red", "blue", "green")[unclass(iris$Species)])
```

## Pairwise Plots
```{r, echo=FALSE, out.width='100%'}
pairs(as.matrix(iris[,-5]), pch=21, bg=c("red", "blue", "green")[unclass(iris$Species)])
```

## Pairwise Plots
- Since the upper and lower graphs are duplicates of each other, we can augment our code to display the **correlation** between our variables in the upper level.

- The correlation measures the strength of the relationship between two random variables.

- Correlations range from -1 to 1, where:

-- Values near 1 indicate a strong positive relationship

-- Values near -1 indicate a strong negative relationship

-- Values near 0 indicate no relationship.

## Pairwise Plots

```{r, eval=FALSE, echo=TRUE,results=FALSE}
panel.pearson <- function(x, y, ...) {
  horizontal <- (par("usr")[1] + par("usr")[2]) / 2;
  vertical <- (par("usr")[3] + par("usr")[4]) / 2;
  text(horizontal, vertical, format(abs(cor(x,y)), digits=2))}

pairs(as.matrix(iris[1:4]), main = "Iris Data", pch = 21, 
      bg = c("red","blue", "green")[unclass(iris$Species)],
      upper.panel=panel.pearson)
```

## Pairwise Plots

```{r, echo=FALSE, out.width='100%'}
panel.pearson <- function(x, y, ...) {
  horizontal <- (par("usr")[1] + par("usr")[2]) / 2;vertical <- (par("usr")[3] + par("usr")[4]) / 2;
text(horizontal, vertical, format(abs(cor(x,y)), digits=2))}

pairs(as.matrix(iris[1:4]), main = "Iris Data", pch = 21, bg = c("red","blue", "green")[unclass(iris$Species)], upper.panel=panel.pearson)
```

## Histograms 

- A **Histogram** is a plot that breaks the data into bins (or breaks) and shows the frequency distribution of those bins.  

- We can change the breaks to see the effect it has data visualization. 

## Histograms 
Let's create some histograms of our Iris data. The number of bins in the histogram is variable.  

```{r}
hist(iris$Petal.Width, breaks=13)
```

## Histograms

```{r}
hist(iris$Petal.Width, breaks=25)
```

## Histograms
We can create custom break points by defining a sequence vector,  `b`, that ranges from `min(iris$Petal.Width)` to the `max(iris$Petal.Width)` with a specified number of breaks.

```{r, results='hide', eval=FALSE, echo=TRUE}
b <- seq(min(iris$Petal.Width), 
         max(iris$Petal.Width), length=11)
b
hist(iris$Petal.Width, breaks=b, 
     xlab="Petal Width", main="Histogram of Petal Width")
```

## Histograms
```{r, echo=FALSE, out.width='65%'}
b <- seq(min(iris$Petal.Width), 
         max(iris$Petal.Width), length=11)

hist(iris$Petal.Width, breaks=b, col="blue", xlab="Petal Width", main="Histogram of Petal Width")
```

## Density Plots
Density plots can be viewed as smoothed versions of a histogram.  We can estimate the density using R's `density` function

```{r}
dens.pw = density(iris$Petal.Width)
plot(dens.pw, ylab = "Frequency", xlab = "Width", 
     main= "Petal Width Density")
```

## Density Plots

```{r}
dens.pl = density(iris$Petal.Length)
plot(dens.pl, ylab = "Frequency", xlab = "Length", 
     main= "Petal Length Density")
```

## Density Plots

```{r}
dens.sw = density(iris$Sepal.Width)
plot(dens.sw, ylab = "Frequency", xlab = "Width", 
     main= "Sepal Width Density")
```
 
## Density Plots

```{r}
dens.sl = density(iris$Sepal.Length)
plot(dens.sl, ylab = "Frequency", xlab = "Length", 
     main= "Sepal Length Density")
```
 
## Density Plots
Let's also look at the density function of `Petal.Length` for each of the three classes of irises.

```{r, echo=FALSE}
den.set = density(iris$Petal.Length
                         [iris$Species == "setosa"])
den.ver = density(iris$Petal.Length
                             [iris$Species == "versicolor"])
den.vir = density(iris$Petal.Length
                            [iris$Species == "virginica"])
```

```{r,results='hide', eval=FALSE, echo=FALSE}
plot(den.set, ylab="Frequency", xlab="Length (cm)",
     main="Density Plot of Petal Lengths", 
     xlim = c(0,8), lwd=2, col="red")
lines(den.ver, col="blue", lwd=2)
lines(den.vir, col="green", lwd=2)
legend(4, 2.75, c("setosa", "versicolor", "virginica"), 
       lwd=rep(2,1), 
       col=c("red", "blue", "green"))
```

## Density Plots

```{r,echo=FALSE,out.width='100%'}
plot(den.set, ylab="Frequency", xlab="Length (cm)",
     main="Density plot of Petal Lengths", 
     xlim = c(0,8), lwd=2, col="red")
lines(den.ver, col="blue", lwd=2)
lines(den.vir, col="green", lwd=2)
legend(4, 2.85, c("Setosa", "Versicolor", "Virginica"), 
       lwd=rep(2,1), 
       col=c("red", "blue", "green"))
```


## Quantile Plots
We can calculate the quantiles of the iris dataset to compare them to those of a normal distribution. 

```{r, results='hide', eval=TRUE, echo=FALSE}
ps = seq(0,1,length=25)
quantile.setosa = quantile(iris$Petal.Length[iris$Species == "setosa"], probs=ps, xlab = "Normal Distrubution Quantile")

quantile.versicolor = quantile(iris$Petal.Length[iris$Species == "versicolor"], probs=ps)

quantile.virginica = quantile(iris$Petal.Length[iris$Species == "virginica"], probs=ps)
```

```{r}
qqnorm(quantile.virginica, main="Virginica")
qqline(quantile.virginica)
```

## Quantile Plots
```{r}
qqnorm(quantile.versicolor, main="Versicolor")
qqline(quantile.versicolor)
```

## Quantile Plots
```{r}
qqnorm(quantile.setosa, main="Setosa")
qqline(quantile.setosa)
```

## Contour Plots
Density estimation is available for higher dimensional data using Contour plots.

- A **contour plot** is a graph that explores the potential relationship among three variables. 

- Contour plots display the 3-dimensional relationship in two dimensions, with x and y variables plotted on the x and y scales and the z variable represented by contours.

- A contour plot is like a topographical map in which x, y, and z values are plotted instead of longitude, latitude, and elevation. 

## Contour Plots
```{r}
library(MASS)
petal.dens = kde2d(iris$Petal.Length, iris$Petal.Width)
contour(petal.dens)
```

## Contour Plots
The plot may also be viewed as a heatmap, with brighter colors denoting higher values.
```{r}
image(petal.dens)
```
